A Decision Procedure for String Constraints with String-Integer Conversion and Flat Regular Constraints

### 2021-09-17 FMCAD2021/1_intro.tex
.

Aiming at introducing Point's procedure to computer science researchers, we reformulate (and slightly improve) Point's procedure in a way, which, hopefully, is more accessible for computer science researchers (Section~\ref{sec-dec}). This can be seen as another contribution of this paper.
%
Furthermore, we propose various optimizations (Section~\ref{sec-opt}) and achieve the first prototypical implementation of Point's procedure (Section~\ref{sec-exp}).

%
%The optimizations include 
%over approximation to reduce the number of subcases
% the elimination of linear occurrences of variables by under approximation and synchronize the elimination of exponential occurrences of the same variable in different atomic formulas. We also apply an over approximation to reduce the number of subcases and adopt DFS with preprocessing to find possible solutions.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\hide{
Although the general framework of our algorithm is quite akin to Point's,  it contains the following improvements and corrections: 
\begin{itemize}
	\item[i)] The syntax of Presburger arithmetic with exponential functions in our setting is more natural and easy to understand; Point's algorithm needs to transform the formula into DNF (disjunction normal form) before  eliminating any variables, that could increase the length of the considered $\paexp$ formula exponentially. In contrast, our algorithm just assumes that formulas are NNF (negation normal form). Therefore, our algorithm is much easy to follow and more efficient.
\item[ii)] Unlike in Point's algorithm to eliminate the exponential part and then the linear part of the largest variable according to the given order one by one, in contrast, our algorithm eliminates the exponential parts of 
	all variables according to the order and then the resulted linear part, that improve the efficiency very much.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%	
%Moreover, we propose various optimizations for satisfiability problem of quantifier-free $\paexp$ formula for an efficient implementation. 
Two important optimizations are that we avoid the elimination of linear occurrences of variables by under approximation and synchronize the elimination of exponential occurrences of the same variable in different atomic formulas. We also apply an over approximation to reduce the number of subcases and adopt DFS with preprocessing to find possible solutions.

\item[iii)] Additionally, a sub-case of eliminating exponential terms in Point's algorithm is incorrect, we correct it by Lemma \ref{lem:exp-ineq} (see later).
\end{itemize}
 }
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

>>>>

Aiming at introducing Point's procedure to computer science researchers, we reformulate (and slightly improve) Point's procedure in a way, which, hopefully, is more accessible for computer science researchers (Section~\ref{sec-dec}). This can be seen as another contribution of this paper.
\wuhao{We also analyze the deterministic upper bound of the complexity for the first time, the time complexity of eliminating one quantifier is shown to be triply exponential (Appendix~\ref{app:cpx}).} 
Furthermore, we propose various optimizations (Section~\ref{sec-opt}) and achieve the first prototypical implementation of Point's procedure (Section~\ref{sec-exp}).

### 2021-09-17 FMCAD2021/paper.tex
complexity done
.

\usepackage{graphicx}
\usepackage[ruled,vlined]{algorithm2e}
\usepackage{amsfonts}
\usepackage{verbatim}
\usepackage{amsmath}
\usepackage{times} 
\usepackage{tikz}
\usepackage{multirow}

\usetikzlibrary{automata, positioning, arrows}
\tikzset{
->, % makes the edges directed
>=stealth, % makes the arrow heads bold
node distance=2cm, % specifies the minimum distance between two nodes. Change if necessary.
every state/.style={thick, fill=gray!10}, % sets the properties for each ’state’ node
initial text=$ $, % sets the text that appears on the start arrow
}

@@

\begin{abstract}
    String constraint solving is the core of various testing and verification approaches for scripting languages. 
    Among algorithms for solving string constraints, flattening is a well-known approach that is particularly useful in handling satisfiable instances.
    As string-integer conversion is an important function appearing in almost all scripting languages, Abdulla et al. extended the flattening approach to this function recently.
    However, their approach supports only a special flattening pattern and leaves the support of the general flat regular constraints as an open problem.
    In this paper, we fill the gap and propose a complete flattening approach for the string-integer conversion. The approach is built upon a quantifier elimination procedure for the linear-exponential arithmetic constraints (namely, the extension of Presburger arithmetic with exponential functions) proposed by Point in 1986. \wuhao{We analyze the complexity of the algorithm and give a triply exponential upper bound both in time and space.} While the quantifier elimination procedure by Point is too expensive to be implemented efficiently, we propose various optimizations and provide a prototypical implementation.
    We evaluate the performance of our implementation on the benchmarks that are generated from the string hash functions as well as randomly.
    The experimental results show that our implementation outperforms the state-of-the-art solvers.

>>>>

\usepackage{graphicx}
\usepackage[ruled,vlined]{algorithm2e}
\usepackage{amsfonts}
\usepackage{verbatim}
\usepackage{amsmath}
\usepackage{times} 
\usepackage{tikz}
\usepackage{multirow}
\usepackage{amssymb}
\usetikzlibrary{automata, positioning, arrows}
\tikzset{
->, % makes the edges directed
>=stealth, % makes the arrow heads bold
node distance=2cm, % specifies the minimum distance between two nodes. Change if necessary.
every state/.style={thick, fill=gray!10}, % sets the properties for each ’state’ node
initial text=$ $, % sets the text that appears on the start arrow
}

@@

\begin{abstract}
    String constraint solving is the core of various testing and verification approaches for scripting languages. 
    Among algorithms for solving string constraints, flattening is a well-known approach that is particularly useful in handling satisfiable instances.
    As string-integer conversion is an important function appearing in almost all scripting languages, Abdulla et al. extended the flattening approach to this function recently.
    However, their approach supports only a special flattening pattern and leaves the support of the general flat regular constraints as an open problem.
    In this paper, we fill the gap and propose a complete flattening approach for the string-integer conversion. The approach is built upon a quantifier elimination procedure for the linear-exponential arithmetic constraints (namely, the extension of Presburger arithmetic with exponential functions) proposed by Point in 1986. \wuhao{The complexity of the algorithm is analyzed for the first time and we show that the elimination of one existential quantifier is 3-EXPSPACE.} While the quantifier elimination procedure by Point is too expensive to be implemented efficiently, we propose various optimizations and provide a prototypical implementation.
    We evaluate the performance of our implementation on the benchmarks that are generated from the string hash functions as well as randomly.
    The experimental results show that our implementation outperforms the state-of-the-art solvers.
